---
title: "ScreenPassTrend"
output: html_document
date: "2023-09-15"
---
```{r}
library(tidyverse)
library(ggplot2)
```


```{r}
data2023 <- nflreadr::load_pbp(2023) 
ftndata2023 <- nflreadr::load_ftn_charting(2023)

data2022 <- nflreadr::load_pbp(2022) 
ftndata2022 <- nflreadr::load_ftn_charting(2022)
```

How to join? 
```{r}
range(data2023$play_id)

range(ftndata2023$nflverse_play_id)

data2023 %>%
  filter(play_id < 39) %>%
  select(game_seconds_remaining) %>%
  range()
```

Ranges are close for play_ids, but not exact. Looking more into it, there are only 32 total rows that have a play_id less than 39, indicating they are outliers of some sort. Seeing their range, its all the first rows of each game, which is probably just the initial kickoff

```{r}
screen_passes_2023 = 
  data2023 %>%
  left_join(ftndata2023, by = c("game_id" = "nflverse_game_id", "play_id" = "nflverse_play_id")) %>%
  filter(is_screen_pass == TRUE) %>%
  select(game_id, week.x, posteam, defteam, qtr, quarter_seconds_remaining, down, ydstogo, ydsnet, yardline_100, 
         desc, 
         yards_gained, starting_hash, pass_location, shotgun, air_yards, yards_after_catch, 
         posteam_score, defteam_score, 
         passer_player_name, receiver_player_name,
         td_prob, ep, epa, total_home_epa, total_away_epa, wp, def_wp, wpa, 
         solo_tackle, pass_attempt, touchdown, complete_pass, 
         cp, cpoe, series_success, 
         series_result, drive_start_yard_line, drive_end_yard_line, aborted_play, success, 
         qb_epa, xyac_epa, xyac_mean_yardage)

screen_passes_2022 = 
  data2022 %>%
  left_join(ftndata2022, by = c("game_id" = "nflverse_game_id", "play_id" = "nflverse_play_id")) %>%
  filter(is_screen_pass == TRUE) %>%
  select(game_id, week.x, posteam, defteam, qtr, quarter_seconds_remaining, down, ydstogo, ydsnet, yardline_100, 
         desc, 
         yards_gained, starting_hash, pass_location, shotgun, air_yards, yards_after_catch, 
         posteam_score, defteam_score, 
         passer_player_name, receiver_player_name,
         td_prob, ep, epa, total_home_epa, total_away_epa, wp, def_wp, wpa, 
         solo_tackle, pass_attempt, touchdown, complete_pass, 
         cp, cpoe, series_success, 
         series_result, drive_start_yard_line, drive_end_yard_line, aborted_play, success, 
         qb_epa, xyac_epa, xyac_mean_yardage) %>%
  mutate(yards_gained = coalesce(yards_gained, 0))

screen_passes_2023_w1 = 
  screen_passes_2023 %>%
  filter(week.x == 1)

screen_passes_2023_w2 = 
  screen_passes_2022 %>%
  filter(week.x == 2)

screen_passes_2022_w1 = 
  screen_passes_2022 %>%
  filter(week.x == 1)
```

```{r}
ggplot(screen_passes_2023, aes(x = yards_gained)) +
  geom_histogram() + 
  labs(title = "Screen Passes Yards Gained 2023",
       subtitle = "Week 1 + 2 Data")

ggplot(screen_passes_2023_w1, aes(x = yards_gained)) +
  geom_histogram() + 
  labs(title = "Screen Passes Yards Gained 2023",
       subtitle = "Week 1 Data")

ggplot(screen_passes_2022, aes(x = yards_gained)) +
  geom_histogram() + 
  labs(title = "Screen Passes Yards Gained 2022",
       subtitle = "All weeks Data")

ggplot(screen_passes_2022_w1, aes(x = yards_gained)) +
  geom_histogram() + 
  labs(title = "Screen Passes Yards Gained 2022",
       subtitle = "Week 1 Data")
```
```{r}
mean(screen_passes_2023$yards_gained)
mean(screen_passes_2023_w1$yards_gained)


mean(screen_passes_2022$yards_gained, na.rm = T)
mean(screen_passes_2022_w1$yards_gained)


data2023
```

Seems like week 1 of 2023 had a lower average than the 2022 season at 4.9 yards (0.2 yards), and an even lower average for week 1 of 2022 at 5.7 yards (0.7 yards). Week 2 rebounded, and raised the average back too 

```{r}
wilcox.test(screen_passes_2023_w1$yards_gained, screen_passes_2022_w1$yards_gained)
```

The difference is not significant to any degree. Let's look at the screen passes that went for negative yards 

```{r}
sum(screen_passes_2023$yards_gained < 0) / length(screen_passes_2023$yards_gained)
sum(screen_passes_2022$yards_gained < 0, na.rm = TRUE) / length(screen_passes_2022$yards_gained)
sum(screen_passes_2022_w1$yards_gained < 0, na.rm = TRUE) / length(screen_passes_2022_w1$yards_gained)

```

Compared to all the screen passes of 2022, the first week of 2023 had about 5% more screen passes go for negative yards

```{r}
screen_passes_2022 %>%
  filter(posteam == 'NE')

screen_passes_2022 %>%
  group_by(posteam) %>%
  summarise(
    num_screens = n(), 
    yds = mean(yards_gained)) %>%
  arrange(-yds)

screen_passes_2022 %>%
  filter(posteam == "TEN")

screen_passes_2022 %>%
  group_by(passer_player_name) %>%
  summarise(
    num_screens = n(), 
    yds = mean(yards_gained)) %>%
  filter(num_screens > 5) %>%
  arrange(-yds)

screen_passes_2022 %>%
  group_by(receiver_player_name) %>%
  summarise(
    num_screens = n(), 
    yds = mean(yards_gained)) %>%
  filter(num_screens > 5) %>%
  arrange(-yds)


screen_passes_2022 %>%
  group_by(defteam) %>%
  summarise(
    num_screens = n(), 
    yds = mean(yards_gained)) %>%
  arrange(-yds)

```
Every team ranged from about 3.7 yards to 6.6 yards per screen pass outside of 3 teams.

Tenessee were extraordinarly good at 8.46 yards per screen, but only ran it 48 times, compared to the 60-70 average

Seattle the New York Jets were all the lowest, at 3.3 and 2.7. 

Pittsburgh was great at defending the screen, with a 1.7 average yards per screen, about 1.4 yards lower than the next team, and teams knew it with only 29 screens run against them, 8 fewer than the next team

```{r}
screen_passes_2023 %>%
  group_by(posteam) %>%
  summarise(
    num_screens = n(), 
    yds = mean(yards_gained)) %>%
  arrange(-yds)

screen_passes_2023 %>%
  filter(posteam == "TEN")

screen_passes_2023 %>%
  filter(posteam == "SEA")

screen_passes_2023 %>%
  group_by(passer_player_name) %>%
  summarise(
    num_screens = n(), 
    yds = mean(yards_gained)) %>%
  filter(num_screens > 1) %>%
  arrange(-yds)

screen_passes_2023 %>%
  group_by(receiver_player_name) %>%
  summarise(
    num_screens = n(), 
    yds = mean(yards_gained)) %>%
  filter(num_screens > 1) %>%
  arrange(-yds)

screen_passes_2023 %>%
  group_by(defteam) %>%
  summarise(
    num_screens = n(), 
    yds = mean(yards_gained)) %>%
  arrange(-yds)
```

And at the start of this year, Tennesee is still at the top at 14 yards per screen. While New York has managed to turn it around, Seattle remains at the bottom, as the only team this season to average negative yards, with their biggest at 2 yards.


Lets check via some other game variables such as quarter, down, yards to go, etc via yards gained, and also EPA (Expected Points Added)

```{r}
screen_passes_2022 %>%
  group_by(qtr) %>%
  summarise(
    count = n(),
    yds = mean(yards_gained),
    epa = mean(epa)
  )

screen_passes_2022 %>%
  group_by(down) %>%
  summarise(
    count = n(),
    yds = mean(yards_gained),
    epa = mean(epa)
  )

screen_passes_2022 %>%
  group_by(ydstogo) %>%
  summarise(
    count = n(),
    yds = mean(yards_gained),
    epa = mean(epa)
  )

## group into 1-3, 4-6, 7-9, 10, 11-15, 16-29

breaks <- c(1, 3, 6, 9, 10, 15, 29, Inf)
labels <- c("Short", "Medium", "Long", "Start", "Start_plus_short", "Start_plus_long", "NA")
screen_passes_2022$ydstogo_group <- cut(screen_passes_2022$ydstogo, breaks = breaks, labels = labels, include.lowest = TRUE)

screen_passes_2022 %>%
  group_by(ydstogo_group) %>%
  summarise(
    count = n(),
    yds = mean(yards_gained),
    epa = mean(epa)
  )

screen_passes_2022 %>%
  group_by(down, ydstogo_group) %>%
  summarise(
    count = n(),
    yds = mean(yards_gained),
    epa = mean(epa)
  ) %>%
  filter(count > 50) %>%
  arrange(-epa)
  

screen_passes_2022 %>%
  group_by(abs(air_yards)) %>%
  summarise(
    count = n(),
    yds = mean(yards_gained),
    epa = mean(epa)
  )


## group by 20s

breaks <- c(1, 20, 40, 60, 80, 100)
labels <- c("OwnEndzone", "OwnMid", "Midfield", "OppMid", "Redzone")
screen_passes_2022$area_of_field <- cut(screen_passes_2022$yardline_100, breaks = breaks, labels = labels, include.lowest = TRUE)


screen_passes_2022 %>%
  group_by(area_of_field) %>%
  summarise(
    count = n(),
    yds = mean(yards_gained),
    epa = mean(epa)
  )
```

- Quarter did not really matter, but Q2 had a closer to 0 compared to quarters 1, 3, and 4
--- Actually higher that the average EPA for all plays
- 2nd down was considerably higher, at 0.01 compared to -0.08
- Medium down (4-6 yards to go) plays, had the highest EPA, next being 11-15 yards to go (+0.03 vs -0.02)
- Any second down scenario, but specifically 2nd and anything more than 4, were extremely beneficial compared to the rest, at the only positive rates. 
- shorter air yard screens (1-2 yards) were generally better, but not by much, especially considering screens with 0 air yards were as bad as the rest
- Screen passes most frequently happen between the opponents 40 and the redzone, but are less ineffective EPA wise when its past their own 40 yard line.


EPA understanding and reference
https://www.fantasylife.com/articles/redraft/what-is-epa-in-fantasy-football-and-sports-betting#:~:text=Expected%20points%20(EP)%20are%20values,one%20play%20to%20the%20next.
```{r}
mean(screen_passes_2022$epa, na.rm = T)
mean(data2022$epa, na.rm = T)
```



Now lets look into the play's positioning 
- QB's stance (shotgun or not)
- Hash vs pass location (same or different or middle)

```{r}
screen_passes_2022 %>%
  group_by(shotgun) %>%
  summarise(
    count = n(),
    yds = mean(yards_gained),
    epa = mean(epa)
  )

screen_passes_2022$starting_hash = ifelse(screen_passes_2022$starting_hash == "L", "left", 
                                          ifelse(screen_passes_2022$starting_hash == "R", "right", "middle"))
screen_passes_2022$net_direction =
  ifelse(screen_passes_2022$starting_hash == "middle", "middle_side",
         ifelse(screen_passes_2022$starting_hash == screen_passes_2022$pass_location, "same_side", "opposite_side"))

screen_passes_2022 %>%
  group_by(net_direction) %>%
  summarise(
    count = n(),
    yds = mean(yards_gained),
    epa = mean(epa)
  )

```

- Most screen passes were run in shotgun, so even though there was a higher yardage for non-shotgun, sample size is relatively insignificant
- most screen passes were ran from opposite sides, so if the ball started near or at the left hash, the screen was ran to the right (nearly 60% of all screen passes, as opposite to the same side being only about 30%) 
---- yardage was relatively similar, with same_side being higher by about 0.45 yards, and +0.04 epa




