---
title: "BengalsStruggles"
output: html_document
date: "2023-10-04"
---

```{r}
library(tidyverse)
library(ggrepel)
library(nflreadr)
library(nflplotR)
```

```{r}
data2022 <- nflreadr::load_pbp(2022) 
ftndata2022 <- nflreadr::load_ftn_charting(2022)

data2023 <- nflreadr::load_pbp(2023) 
ftndata2023 <- nflreadr::load_ftn_charting(2023)
```


```{r}
#Filter out "no plays"
data2022$no_play = ifelse(str_detect(data2022$desc, "No Play"), 1, 0)
data2023$no_play = ifelse(str_detect(data2023$desc, "No Play"), 1, 0)
```

```{r}
data2023 = 
  data2023 %>%
  left_join(ftndata2023, by = c("game_id" = "nflverse_game_id", "play_id" = "nflverse_play_id")) %>%
  #filter for only weeks 1-4
  filter(week.x %in% c(1:4)) %>%
  #filter out plays where offensive team committed a penalty (no play)
  filter(!(!is.na(penalty_team) & (penalty_team == posteam))) %>%
  filter(no_play == 0) %>%
  #filter out non-plays 
  filter(play_type %in% c('pass', 'run'))

data2022 = 
  data2022 %>%
  left_join(ftndata2022, by = c("game_id" = "nflverse_game_id", "play_id" = "nflverse_play_id")) %>%
  #filter for only regular season
  filter(season_type == "REG") %>%
  #filter out plays where offensive team committed a penalty (no play)
  filter(!(!is.na(penalty_team) & (penalty_team == posteam))) %>%
  filter(no_play == 0) %>%
  #filter out non-plays 
  filter(play_type %in% c('pass', 'run'))

```


```{r}
cin_data2023 = 
  data2023 %>%
  filter(posteam == "CIN") %>%
  group_by(shotgun) %>%
  summarise(
    count = n(), 
    perc = n() / nrow(data2023 %>% filter(posteam == "CIN")),
    yds = mean(yards_gained, na.rm = T),
    epa = mean(epa, na.rm = T)
  ) %>%
  mutate(year = 2023)




cin_data2022 = 
  data2022 %>%
  filter(posteam == "CIN") %>%
  group_by(shotgun) %>%
  summarise(
    count = n(), 
    perc = n() / nrow(data2022 %>% filter(posteam == "CIN")),
    yds = mean(yards_gained, na.rm = T),
    epa = mean(epa, na.rm = T)
  ) %>%
  mutate(year = 2022)


data2022 %>%
  filter(posteam == "CIN") %>%
  filter(play_type == "run") %>%
  group_by(shotgun, is_play_action) %>%
  summarise(
    count = n(), 
    perc = n() / nrow(data2022 %>% filter(posteam == "CIN")),
    yds = mean(yards_gained, na.rm = T),
    epa = mean(epa, na.rm = T)
  ) %>%
  mutate(year = 2022)
```

```{r}
rbind(cin_data2022, cin_data2023) %>%
  mutate(shotgun = as.factor(shotgun),
         year = as.factor(year)) %>%
  ggplot(aes(x = shotgun, y = count, fill = year)) + 
  geom_bar(stat="identity", position="fill") + 
  scale_fill_manual(values=c('#FB4F14','#000000')) + 
  labs(
    title = "Bengals Plays via Shotgun",
    y = "Percentage of Plays"
  ) +
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.background = element_blank(), 
    axis.line = element_line(colour = "black"))


cin_data2023[1,5] <- NA

rbind(cin_data2022, cin_data2023) %>%
  mutate(shotgun = as.factor(shotgun), year = as.factor(year)) %>%
  ggplot(aes(x = year)) +
  geom_bar(aes(y = perc, fill = shotgun), stat="identity", position="dodge") + 
  scale_fill_manual(
    values=c('#FFD580','#454545')) +
  geom_point(aes(y = epa+0.2, fill = shotgun, color = factor(epa > 0)), size = 5, shape = 'diamond', position = position_dodge(width = 0.9)) +
  geom_text(aes(y = epa + 0.2 + 0.07, label = round(epa, 2), group = shotgun, color = factor(epa > 0)), size = 5, position = position_dodge(width = 0.9)) +
  scale_color_manual(values = c("red", "lightgreen")) +  # Specify red and green colors
  scale_y_continuous(
    name = "Percentage of Plays",
    sec.axis = sec_axis(~.-0.2, name = "EPA")
  ) +
  labs(
    title = "Bengals Plays via Shotgun",
    subtitle = "With nearly all plays run from shotgun in 2023 (W1-4), Bengals lost their EPA firepower,
                EPA in non-shotgun 2023 is not shown, with only 2 plays ran"
  ) +
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.background = element_blank(), 
    axis.line = element_line(colour = "black")) + 
  guides(color = FALSE)

ggsave("2023BengalsViaShotgun.png")

```

% Plays Run in Shotgun vs Team EPA
```{r}
team_percshotgun_2023 = 
  data2023 %>%
  group_by(posteam, shotgun) %>%
  summarise(
    count = n(), 
    yds = mean(yards_gained, na.rm = T),
    epa = mean(epa, na.rm = T)
  ) %>%
  group_by(posteam) %>%
  mutate(Proportion = count / sum(count)) %>%
  select(posteam, shotgun, Proportion) %>%
  filter(shotgun == 1)

team_epa_2023 = data2023 %>%
  group_by(posteam) %>%
  summarise(
    count = n(), 
    yds = mean(yards_gained, na.rm = T),
    epa = mean(epa, na.rm = T)
  ) %>%
  select(posteam, epa)


team_percshotgun_2023 %>%
  left_join(team_epa_2023) %>%
  ggplot(aes(Proportion, epa)) +
  geom_point(shape = NA) + 
  #add points for the QBs with the logos (this uses nflplotR package)
  geom_nfl_logos(aes(team_abbr = posteam), width = 0.05) + 
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "2023 Plays Ran in Shotgun vs Team Overall EPA",
    subtitle = "Of the 7 teams who have ran the most shotgun in 2023, 5 significant below 0 EPA",
    caption = "2023 Data, Data: @nflfastr",
    x = "Proportion of Plays Ran in Shotgun",
    y = "Team Overall EPA"
  ) +
  theme_classic()

ggsave("2023TeamShotgunVsEPA.png")

```


```{r}
team_percshotgun_2022 = data2022 %>%
  group_by(posteam, shotgun) %>%
  summarise(
    count = n(), 
    yds = mean(yards_gained, na.rm = T),
    epa = mean(epa, na.rm = T)
  ) %>%
  group_by(posteam) %>%
  mutate(Proportion = count / sum(count)) %>%
  select(posteam, shotgun, Proportion) %>%
  filter(shotgun == 1)

team_epa_2022 = data2022 %>%
  group_by(posteam) %>%
  summarise(
    count = n(), 
    yds = mean(yards_gained, na.rm = T),
    epa = mean(epa, na.rm = T)
  ) %>%
  select(posteam, epa)


team_percshotgun_2022 %>%
  left_join(team_epa_2022) %>%
  ggplot(aes(Proportion, epa)) +
  geom_point(shape = NA) + 
  #add points for the QBs with the logos (this uses nflplotR package)
  geom_nfl_logos(aes(team_abbr = posteam), width = 0.05) + 
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "2022 Plays Ran in Shotgun vs Team Overall EPA",
    subtitle = "Of the 9 teams who ran the most shotgun, 7 above 0 EPA including top 4 in the league",
    caption = "2022 Data, Data: @nflfastr",
    x = "Proportion of Plays Ran in Shotgun",
    y = "Team Overall EPA"
  ) +
  theme_classic()

ggsave("2022TeamShotgunVsEPA.png")

```

Previous 3 plays
```{r}

data2022 %>%
  filter(posteam == "CIN") %>%
  mutate(
    lagged_shotgun = ifelse(lag(defteam) != defteam, NA, lag(shotgun)),
    lagged_yds = ifelse(lag(defteam) != defteam, NA, lag(yards_gained)),
    lagged_shotgun_2 = ifelse(lag(defteam) != defteam, NA, lag(shotgun, 2)),
    lagged_yds_2 = ifelse(lag(defteam) != defteam, NA, lag(yards_gained, 2))
  ) %>%
  select(game_id, away_team, shotgun, yards_gained, epa, lagged_shotgun, lagged_yds, lagged_shotgun_2, lagged_yds_2) %>%
  group_by(shotgun, lagged_shotgun, lagged_shotgun_2) %>%
  summarise(
    count = n(),
    yds_gained = mean(yards_gained),
    epa = mean(epa)
  ) %>%
  filter(count > 30) %>%
  arrange(-epa)

data2022 %>%
  mutate(
    lagged_shotgun = ifelse(lag(defteam) != defteam, NA, lag(shotgun)),
    lagged_yds = ifelse(lag(defteam) != defteam, NA, lag(yards_gained)),
    lagged_shotgun_2 = ifelse(lag(defteam) != defteam, NA, lag(shotgun, 2)),
    lagged_yds_2 = ifelse(lag(defteam) != defteam, NA, lag(yards_gained, 2))
  ) %>%
  select(game_id, away_team, shotgun, yards_gained, epa, lagged_shotgun, lagged_yds, lagged_shotgun_2, lagged_yds_2) %>%
  group_by(shotgun, lagged_shotgun, lagged_shotgun_2) %>%
  summarise(
    count = n(),
    yds_gained = mean(yards_gained),
    epa = mean(epa)
  ) %>%
  filter(count > 30) %>%
  arrange(-epa)

data2022 %>%
  mutate(
    lagged_shotgun = ifelse(lag(defteam) != defteam, NA, lag(shotgun)),
    lagged_yds = ifelse(lag(defteam) != defteam, NA, lag(yards_gained)),
    lagged_shotgun_2 = ifelse(lag(defteam) != defteam, NA, lag(shotgun, 2)),
    lagged_yds_2 = ifelse(lag(defteam) != defteam, NA, lag(yards_gained, 2))
  ) %>%
  select(game_id, away_team, desc, play_type, shotgun, yards_gained, epa, lagged_shotgun, lagged_yds, lagged_shotgun_2, lagged_yds_2) %>%
  filter(
    (shotgun == 0) & (lagged_shotgun == 1) & (lagged_shotgun_2 == 1)
  ) %>%
  arrange(-epa) %>%
  ggplot(aes(epa, color = play_type)) +
  geom_density() + 
  geom_vline(xintercept = 0) +
  theme_classic()
  
  
data2022 %>%
 mutate(
    lagged_shotgun = ifelse(lag(defteam) != defteam, NA, lag(shotgun)),
    lagged_yds = ifelse(lag(defteam) != defteam, NA, lag(yards_gained)),
    lagged_shotgun_2 = ifelse(lag(defteam) != defteam, NA, lag(shotgun, 2)),
    lagged_yds_2 = ifelse(lag(defteam) != defteam, NA, lag(yards_gained, 2)),
    lagged_shotgun_3 = ifelse(lag(defteam) != defteam, NA, lag(shotgun, 3)),
    lagged_yds_3 = ifelse(lag(defteam) != defteam, NA, lag(yards_gained, 3))
  ) %>%
  select(game_id, away_team, desc, play_type, shotgun, yards_gained, epa, lagged_shotgun, lagged_yds, lagged_shotgun_2, lagged_yds_2, lagged_shotgun_3, lagged_yds_3) %>%
  #create category for non-shotgun play mixed into last 3 plays
  mutate(
    non_shotgun_last3 = as.factor(ifelse(((lagged_shotgun_3 == 0) | (lagged_shotgun == 0) | (lagged_shotgun_2 == 0)), 1, 0))
  ) %>%
  group_by(non_shotgun_last3) %>%
  summarise(
    count = n(),
    std_epa = sd(epa),
    yds_gained = mean(yards_gained),
    epa = mean(epa)
  )

data2022 %>%
  mutate(
    lagged_shotgun = ifelse(lag(defteam) != defteam, NA, lag(shotgun)),
    lagged_yds = ifelse(lag(defteam) != defteam, NA, lag(yards_gained)),
    lagged_shotgun_2 = ifelse(lag(defteam) != defteam, NA, lag(shotgun, 2)),
    lagged_yds_2 = ifelse(lag(defteam) != defteam, NA, lag(yards_gained, 2)),
    lagged_shotgun_3 = ifelse(lag(defteam) != defteam, NA, lag(shotgun, 3)),
    lagged_yds_3 = ifelse(lag(defteam) != defteam, NA, lag(yards_gained, 3))
  ) %>%
  select(game_id, away_team, desc, play_type, shotgun, yards_gained, epa, lagged_shotgun, lagged_yds, lagged_shotgun_2, lagged_yds_2, lagged_shotgun_3, lagged_yds_3) %>%
  #create category for non-shotgun play mixed into last 3 plays
  mutate(
    non_shotgun_last3 = as.factor(ifelse(((lagged_shotgun_3 == 0) | (lagged_shotgun == 0) | (lagged_shotgun_2 == 0)), 1, 0))
  ) %>%
  filter(!is.na(non_shotgun_last3) & !is.na(epa)) %>%
  ggplot(aes(epa, color = non_shotgun_last3)) +
  geom_density() + 
  geom_vline(xintercept = 0) + 
  labs(
    title = "Running a non-shotgun play within Last 3 Plays Boosts EPA",
    subtitle = "EPA is graphed for Play following L3",
    caption = "2022 Data, Data: @nflfastr",
    color = "Non-Shotgun Ran L3 Plays"
  ) +
  theme_classic()

ggsave("2022ShotgunL3.png")
  
```

Bengals ran shotgun in 3 consecutive plays 54% of the time last year, and was the highest average of shotgun/non-shotgun at 6.4 yards per play and 0.16 EPA

Across the league, when teams ran shotgun twice then non-shotgun was the highest EPA but lower yards gained, presuming that the two shotgun plays set up a great EPA play (first down, shorter TD score)

Running a non-shotgun play within your last 3 were much more effective via EPA, 0.015 epa vs -0.0075 epa and a p-value of 0.17, which accounts for the double in sample size that no shotgun for last 3 has (20k to 10k). And the graph shows it to be pretty agnostic of play type with identical graphs, but the red line (a non-shotgun play ran in last 3) edging the high EPA range of 2-5. 


Big plays
```{r}
data2022 %>%
  filter(posteam == "CIN") %>%
  mutate(
    lagged_shotgun = ifelse(lag(defteam) != defteam, NA, lag(shotgun)),
    lagged_yds = ifelse(lag(defteam) != defteam, NA, lag(yards_gained)),
    lagged_shotgun_2 = ifelse(lag(defteam) != defteam, NA, lag(shotgun, 2)),
    lagged_yds_2 = ifelse(lag(defteam) != defteam, NA, lag(yards_gained, 2))
  ) %>%
  select(game_id, away_team, shotgun, yards_gained, epa, lagged_shotgun, lagged_yds, lagged_shotgun_2, lagged_yds_2) %>%
  filter(yards_gained >= 15) %>%
  group_by(shotgun, lagged_shotgun, lagged_shotgun_2) %>%
  summarise(
    count = n(),
    yds_gained = mean(yards_gained),
    epa = mean(epa)
  ) %>%
  arrange(-epa)

data2022 %>%
  mutate(
    lagged_shotgun = ifelse(lag(defteam) != defteam, NA, lag(shotgun)),
    lagged_yds = ifelse(lag(defteam) != defteam, NA, lag(yards_gained)),
    lagged_shotgun_2 = ifelse(lag(defteam) != defteam, NA, lag(shotgun, 2)),
    lagged_yds_2 = ifelse(lag(defteam) != defteam, NA, lag(yards_gained, 2))
  ) %>%
  select(game_id, away_team, shotgun, yards_gained, epa, lagged_shotgun, lagged_yds, lagged_shotgun_2, lagged_yds_2) %>%
  filter(yards_gained >= 15) %>%
  group_by(shotgun, lagged_shotgun, lagged_shotgun_2) %>%
  summarise(
    count = n(),
    yds_gained = mean(yards_gained),
    epa = mean(epa)
  ) %>%
  arrange(-epa)
```
For the entire league, when looking at plays that gained over 15 yards, the most effective EPA combination of plays was running a non-shotgun play within your previous 2 plays, then followed by a shotgun play, at a 2.25+ EPA on average. Not neccesarily an indicator of what makes the big plays, as it could've been based on the situation  








Week 5 Update after Bengals Win:

```{r}
data2023 <- nflreadr::load_pbp(2023) 
ftndata2023 <- nflreadr::load_ftn_charting(2023)
```


```{r}
#Filter out "no plays"
data2023$no_play = ifelse(str_detect(data2023$desc, "No Play"), 1, 0)
```

```{r}
data2023 = 
  data2023 %>%
  left_join(ftndata2023, by = c("game_id" = "nflverse_game_id", "play_id" = "nflverse_play_id")) %>%
  #filter out plays where offensive team committed a penalty (no play)
  filter(!(!is.na(penalty_team) & (penalty_team == posteam))) %>%
  filter(no_play == 0) %>%
  #filter out non-plays 
  filter(play_type %in% c('pass', 'run'))

```

```{r}
cin_week5 = data2023 %>%
  left_join(ftndata2023, by = c("game_id" = "nflverse_game_id", "play_id" = "nflverse_play_id")) %>%
  #filter for only weeks 1-4
  filter(week.x %in% c(5)) %>%
  #filter out plays where offensive team committed a penalty (no play)
  filter(!(!is.na(penalty_team) & (penalty_team == posteam))) %>%
  filter(no_play == 0) %>%
  #filter out non-plays 
  filter(play_type %in% c('pass', 'run')) %>%
  #only CIN
  filter(posteam == "CIN")

cin_week5 %>%
  group_by(shotgun) %>%
  summarise(
    count = n(), 
    perc = n() / nrow(data2022 %>% filter(posteam == "CIN")),
    yds = mean(yards_gained, na.rm = T),
    epa = mean(epa, na.rm = T)
  )

league = data2023 %>%
  left_join(ftndata2023, by = c("game_id" = "nflverse_game_id", "play_id" = "nflverse_play_id")) %>%
  #filter out plays where offensive team committed a penalty (no play)
  filter(!(!is.na(penalty_team) & (penalty_team == posteam))) %>%
  filter(no_play == 0) %>%
  #filter out non-plays 
  filter(play_type %in% c('pass', 'run'))

league %>%
  group_by(shotgun) %>%
  summarise(
    count = n(), 
    perc = n() / nrow(data2022 %>% filter(posteam == "CIN")),
    yds = mean(yards_gained, na.rm = T),
    epa = mean(epa, na.rm = T)
  )
```

```{r}
team_percshotgun_2023 = data2023 %>%
  filter(week.x %in% c(5)) %>%
  group_by(posteam, shotgun) %>%
  summarise(
    count = n(), 
    yds = mean(yards_gained, na.rm = T),
    epa = mean(epa, na.rm = T)
  ) %>%
  group_by(posteam) %>%
  mutate(Proportion = count / sum(count)) %>%
  select(posteam, shotgun, Proportion) %>%
  filter(shotgun == 1)



team_epa_2023 = data2023 %>%
  filter(week.x %in% c(5)) %>%
  group_by(posteam) %>%
  summarise(
    count = n(), 
    yds = mean(yards_gained, na.rm = T),
    epa = mean(epa, na.rm = T)
  ) %>%
  select(posteam, epa)


team_percshotgun_2023 %>%
  left_join(team_epa_2023) %>%
  ggplot(aes(Proportion, epa)) +
  geom_point(shape = NA) + 
  #add points for the QBs with the logos (this uses nflplotR package)
  geom_nfl_logos(aes(team_abbr = posteam), width = 0.05) + 
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Week 5 2023 Plays Ran in Shotgun vs Team Overall EPA",
    subtitle = "Bengals Offense performs at league average, with nearly all Plays in Shotgun",
    caption = "Raj Dasani | Data: @nflfastr",
    x = "Proportion of Plays Ran in Shotgun",
    y = "Team Overall EPA"
  ) +
  theme_classic()
```

```{r}
ind_week5 = data2023 %>%
  left_join(ftndata2023, by = c("game_id" = "nflverse_game_id", "play_id" = "nflverse_play_id")) %>%
  #filter for only weeks 1-4
  filter(week.x %in% c(5)) %>%
  #filter out plays where offensive team committed a penalty (no play)
  filter(!(!is.na(penalty_team) & (penalty_team == posteam))) %>%
  filter(no_play == 0) %>%
  #filter out non-plays 
  filter(play_type %in% c('pass', 'run')) %>%
  #only CIN
  filter(posteam == "IND")

ind_week5 %>%
  group_by(shotgun) %>%
  summarise(
    count = n(), 
    perc = n() / nrow(data2022 %>% filter(posteam == "CIN")),
    yds = mean(yards_gained, na.rm = T),
    epa = mean(epa, na.rm = T)
  )
```

